\LARGE \bf Eye tracking with Educational Robots: A Cautionary Tale

### 2014-10-02 paper.tex
en_US -> en_GB + other minor stuff

%%% HRI 2015 -> double-blind review process

%\numberofauthors{1} 
%\author{
%\alignauthor
%Kshitij Sharma, S\'{e}verin Lemaignan, Pierre Dillenbourg\\
%       \affaddr{Computer-Human Interaction in Learning and Instruction Lab (CHILI)}\\
%       \affaddr{Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL)}\\
%       \affaddr{CH-1015 Lausanne}\\
%       \affaddr{Switzerland}\\
%       \email{firstname.lastname@epfl.ch}
%}
%
%\additionalauthors{Additional authors: 
%Francesco Mondada, LSRO, EPFL, francesco.mondada@epfl.ch 
%}
%

@@

We present an eye-tracking study on an educational robot
Thymio II with 52 participants. The participants observed the robot
detecting obstacles in the first phase of the experiment and they
interacted with the robot in the second phase of the experiment.
Finally, they were asked to explain the functionality of the robot. The
gaze of the participants was recorded during both the phases. The values
from the robot's sensors were visualized on the robot's body using LEDs
in two different ways in the two experimental conditions and there was
no sensor data visualization in the control condition. We found that the
sensor data visualization has an impact on the gaze patterns and the
performance of the participants. The main goal of the experiment was to
investigate that eye tracking could bring insights to the effectiveness
of the interaction between the human user and the robot.

@@

On the other hand, educational robots can be of great use in making young school
kids interested in technological developments~\cite{cooper1999robots,
wyffels2010building}. It is important to know about the key socio-cognitive
features of this subset of human-robot interaction because these features let
the researchers design better interactive scenarios in the future. These
features include the conceptual mapping of the sensor data visualization and the
robot's function and the generalizing the concepts to solve a given problem in a
contextual learning environment. ~

Thymio II~\cite{riedo2012two} can provide salient effects onto our visual
attention through the LEDs. We can use this feature to guide the users'
attention in a way that the effectiveness of the interaction is maximized. In
the case of interaction with the educational robots the effect of the
interaction is to make the users understand the technical aspects of the robot
and to make the users be able to transfer it to different domains.

@@

Our working hypothesis is that with careful mapping between the LED
visualizations and the behavior of the robot we can induce the understanding of
the functionality of the robot very easily. We propose to use the gaze data to
verify this hypothesis. With eye-tracking data we can judge whether the mapping
between the behavior and visual actuators is useful or not for the participants'
understanding.  Moreover, we hypothesize that the good performers will be able
to anticipate the robot's motion. Eye-tracking data can also be useful to verify
the anticipation hypothesis.

The experiment consists in two phases. The first phase is the observation phase
during which the participants watch the robot move and detect an obstacle in a
``playground''. The second phase requires the participants to interact with the
robot by placing an obstacle in its path to make it detect the obstacle before
reaching a certain target.  After each phase the participant explain the
behavior of the robot. The sensor data was visualized through LEDs with three
visualization conditions: {\sf TRUE} visualization, {\sf RANDOM} visualization and {\sf NONE}
visualization. Through this experiment we investigate the following research
questions:

@@

The rest of the paper is organized as follows. The second section gives
a brief review of the eye-tracking research done to quantify the
expertise and performance. The third section presents the specificities
of the study. The fourth section lists the main results. The fifth
section discusses the results. Finally, the sixth section concludes the
paper.

@@

Our previous work also show a clear relation between gaze patterns and
task-based performance. In a pair-programming task, \cite{jermann2012effects}
showed that the good performing pairs have more synchronized gaze on different
parts of a program than the bad performing pairs. In a similar task,
\cite{sharma2012gaze} showed that the good performing pairs pay more attention
to the data-flow of the program than the poor performing pairs. Moreover,
\cite{sharma2013understanding} showed that while describing the functionality of
a program the well performing teams had more gaze on the variable modification
parts in the program while poor performing teams have equal distribution of gaze
on different parts of the program during similar phase of the task.  Thus we see
that previous research provides insights about the relationship between the gaze
patterns and the behavioral and task-based performance indicators in diverse
scenarios.

@@

In the experiment, there were 3 sensor data visualization conditions.  For the
{\sf TRUE} visualization condition, the number of illuminated LEDs, on the top of the
robot, is proportional to the intensity measured by the IR-sensor. When the
intensity of the reflected IR crosses a predefined threshold, the robot does a
90° turn to its right. For the two other conditions, the behavior is the same,
simply the display changes. In the {\sf RANDOM} condition, the display shows as if the
robot was detecting the obstacle at random moments, while in the {\sf NONE}
visualization condition nothing is shown on top of the robot. The robot has a
fitting for a pencil that we used for the drawing of its trajectory.

@@

The playground (Figure~\ref{playground}) is designed in order to allow the
participants to get a reference of the previous behavior of the robot, as well
as to allow us to localize the position of the robot, the position of the
obstacles and the position of the gaze on the video recorded from the
eye-tracker. The localization is made possible by the use of unique fiducial
markers on the playground. On the right hand side, the observation phase takes
place and there will be the reference for the black and white obstacles. On the
left hand side, the interaction takes place.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{maze}
    \caption{\small Basic playground for the experiment. The fiducial markers
    are used for automatic localization of the gaze pointers on the observation
    video.}

@@

One important fact to be considered is that the localization of the
robot, the obstacle or the gaze pointer on the observation field for the
participant is not a trivial task. The fact that the participants are
free to move, poses a challenge to the eye-tracking analysis. The
movement of the participant causes change in the observation field of
the participant almost every frame of the eye-tracking video. To cope up
with the changing observation field we decided to put an array of
fiducial markers on the playground. Using the fiducial markers enabled
us to recreate the whole playground for every frame in the video
recorded from the eye-tracker's camera (Figure~\ref{expe}).

@@

\paragraph{Observation phase} The participant observes the Thymio II robot
approaching obstacles. The robot turns as soon as it detects the
obstacle. The participant watches the robot's behavior for a black and a
white obstacle for 5 trials each. For the white, the robot turns
earlier; for the black obstacle the robot turns later. In the final
trial the robot is equipped with a pen to draw its trajectory for both
obstacles with different colors. After drawing, the participant is asked
the question: ``How does the robot work?''

\paragraph{Interaction phase} In the second phase, the participant is asked
to guide the robot to a goal, using a gray obstacle. The participant can put the
obstacle wherever he likes on the left hand side (Figure~\ref{playground}) of
the playground before the robot starts. We designed the gray obstacle to reflect
more infrared light (IR) than the black or the white obstacles therefore the
robot turns even earlier. This leads to a surprising effect. The participant has
only the robot's trajectories as a source of information. Each participant is
given 5 trials to put the obstacle in order to guide the robot to the goal. The
participant then has to answer the same question as in the observation phase,
response to which is considered as a final answer.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{setup}
    \caption{\small Experimental setup for observation phase. The observation
    video is recorded from the video camer in the front of the eye tracker (in
    the lop-left corner).}

@@

We transcribed the responses to perform a rating scheme on the level of
abstraction and correctness. A low level of abstraction consists in
describing the functionality of the robot in the terms of the electric
or programming aspects of the robot, a high level represents describing
the behavior. For the levels of correctness, we categorized the
responses into correct or incorrect response. In TABLE I. we present
sample responses for the different categories.

\begin{table}[ht!]
    \centering
    \footnotesize
    \begin{tabular}{ccp{4.5cm}}
        \toprule
        Correct. & Abstract. & Sample answer \\
        \midrule
        \textit{incorrect} & \textit{low} & ``The robot turns at a given distance in respect to a color, defined in the program.''\\ 
        \midrule
        \textit{incorrect} & \textit{high} & `Gray is the most dangerous, so the
        robot turn really early.''\\ 
        \midrule
        \textit{correct} & \textit{low} & ``The gray obstacle reflects a lot,
        the robot sends light. At a certain threshold the robot turns.'' \\ 
        \midrule
        \textit{correct} & \textit{high} & ``The gray obstacle is shinier, so
        more reflective, and better detectable.''\\ 
        \bottomrule
    \end{tabular}
    \caption{Sample answers for all classes}

@@

    \item The participant identified a wrong cause for the behavior of the robot:
        ``The robot notices the tags on the side''.

    \item The participant held the code responsible for the cause of different
        behavior: ``The robot turns at a given distance in respect to a color,
        defined in the program''.

    \item The participant tried to explain by analogies or did not explain the
        robots behavior at an understandable level: ``Gray is the most
        dangerous, so the robot turns really early''.

@@

The most common form of eye-tracking data aggregation are the fixations.
Fixations depict the periods in the user's observation when the user
attends a relatively small part of the visual stimulus for a relatively
longer period of time. The various fixation detection algorithms are
summarized in~\cite{duchowski2007eye}. Following are the two gaze measures, derived
from the fixations, which we used in the present work.

@@

Average fixation on the reference side: The main area (right hand side
of figure 2) in the observation phase is treated as the reference side
in the interaction phase. The rationale behind keeping the lines drawn
by the robot in the observation phase is for the participants to be able
to refer to the prior behavior of the robot in the interaction phase. We
measured the amount of time a participant looks at the reference side
and averaged this duration for the number of fixations on the reference
side for the particular participant.

@@

We found no bias for age, gender, expertise, student status or condition
in respect to the correctness, abstraction and type of mistake made for
the final answer. Surprisingly there is no correlation between the
visualization condition and the correctness or abstraction of the
answers given. We could not find any anticipation patterns in the
eye-tracking data. We found some other relation between the different
measures. We present the results in following subsections.

@@

The average fixation duration on the robot during observation phase
(Figure~\ref{res1}) is significantly more in {\sf TRUE} and {\sf RANDOM} condition than in
the {\sf NONE} condition ($F[2,49]=3.68$, p = .03). This depicts that the
sensor data visualization has an effect on the users' attention.

@@

The fact that the participants in the {\sf RANDOM} or {\sf NONE} visualization
condition improved more than those with {\sf TRUE} condition is surprising
(Figure 5). There are two plausible explanations. The first is that
participants who see the robot in the {\sf TRUE} condition have a stronger
believe that the robot always behaves the same. They see the LEDs of the
robot turning on and therefore the robot still works. The ones with {\sf NONE}
visualization do not have an indication whether the robot works the same
way as in the observation phase, and start experimenting earlier. The
second is that the participants in {\sf TRUE} condition just put the obstacles
out of detection range, because they could see different display
information and concluded therefore it may behave differently.

The fact that participants have lower average fixation duration on the
reference side in the interaction phase in the {\sf TRUE} condition than the
other two conditions actually supports the explanation that participants
with {\sf TRUE} condition had a higher initial belief in their cognitive model
of how does the robot work (Figure 6). They did not need to look onto
the reference given in observation phase. This also verifies our working
hypothesis that a carefully designed mapping between the robot's
behavior and its visual actuators can help in maximizing the learning
effect during the interaction.

@@

The advantages of mobile eye-tracking come with an additional cost. As
we mentioned earlier, the automatic localization of the gaze pointer on
the visual stimulus is not a trivial task. To do this we have to use a
number of fiducial markers to the playground (Figure 2). We observed
that the fiducial markers cause a distraction to the participants. As a
future work, we are trying to find a suitable replacement to the
fiducial markers.

Another plausible explanation for the participants in {\sf NONE} and {\sf RANDOM} condition
being better performers than the {\sf TRUE} condition during the interaction phase
comes from the transparency of the robot's behavior. The robot's
behavior was not at all transparent in the former two conditions. This might
have caused the participants to put more effort in understanding the behavior.
Therefore, they performed better.

@@

We addressed two research questions in the beginning of the present
work. The results show that the way the sensor data is visualized has an
impact on both the gaze patterns and the performance (Question 2).
However, for the relation between the gaze patterns of the participants
and their performance outcomes (Question 1) we cannot claim a direct
relation. We found a negative correlation between the gaze-patterns and
the first improvement during the interaction phase. We could not find a
relation between the explanation of the participants and their
gaze-patterns. Moreover, we did not find any anticipation patterns in
the gaze data. The lack of evidence for the relation between
participants' explanations and their gaze patterns prevents us to make
any strong claims about the relation between the overall performance and
the gaze patterns.

The fact that we cannot find a relation between the gaze patterns and
all the factors in the performance might have its roots in the way we
designed the interaction phase experiment. As a working hypothesis we
said that there should be clear and carefully designed mapping between
the robots' behavior and the sensor data visualization. Its environment
also affects the behavior of the robot and hence the environment also
should be taken into consideration while designing the
behavior-visualization mapping. We introduced a novelty (the gray
obstacle being more reflective) in the environment during the
interaction phase of the experiment that forced the participants to
revise their hypothesis and hence they failed to understand the robot's
functionality. This also serves as a word of caution in future while
designing such interactions. The observation phase serves as a training
session for the participants, which can later be tested during the
interaction phase. Any novel behavior modality (in our case, the more
reflective gray obstacle) can have a negative impact on the
participants' understanding.

>>>>

%%% HRI 2015 -> double-blind review process
%
%\numberofauthors{1} 
%\author{
%\alignauthor
%Kshitij Sharma, S\'{e}verin Lemaignan, Pierre Dillenbourg\\
%       \affaddr{Computer-Human Interaction in Learning and Instruction Lab (CHILI)}\\
%       \affaddr{Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL)}\\
%       \affaddr{CH-1015 Lausanne}\\
%       \affaddr{Switzerland}\\
%       \email{firstname.lastname@epfl.ch}
%}
%
%\additionalauthors{Additional authors: 
%Francesco Mondada, LSRO, EPFL, francesco.mondada@epfl.ch 
%}
%

@@

We present an eye-tracking study on an educational robot
Thymio II~\cite{riedo2012two} with 52 participants. The participants observed the robot
detecting obstacles in the first phase of the experiment and they
interacted with the robot in the second phase of the experiment.
Finally, they were asked to explain the functionality of the robot. The
gaze of the participants was recorded during both the phases. The values
from the robot's sensors were visualised on the robot's body using LEDs
in two different ways in the two experimental conditions and there was
no sensor data visualisation in the control condition. We found that the
sensor data visualisation has an impact on the gaze patterns and the
performance of the participants. The main goal of the experiment was to
investigate that eye tracking could bring insights to the effectiveness
of the interaction between the human user and the robot.

@@

On the other hand, educational robots can be of great use in making young school
kids interested in technological developments~\cite{cooper1999robots,
wyffels2010building}. It is important to know about the key socio-cognitive
features of this subset of human-robot interaction because these features let
the researchers design better interactive scenarios in the future. These
features include the conceptual mapping of the sensor data visualisation and the
robot's function and the generalising the concepts to solve a given problem in a
contextual learning environment. ~

Thymio II~\cite{riedo2012two} can provide salient effects onto our visual
attention through the LEDs. We can use this feature to guide the users'
attention in a way that the effectiveness of the interaction is maximised. In
the case of interaction with the educational robots the effect of the
interaction is to make the users understand the technical aspects of the robot
and to make the users be able to transfer it to different domains.

@@

Our working hypothesis is that with careful mapping between the LED
visualisations and the behaviour of the robot we can induce the understanding of
the functionality of the robot very easily. We propose to use the gaze data to
verify this hypothesis. With eye-tracking data we can judge whether the mapping
between the behaviour and visual actuators is useful or not for the participants'
understanding.  Moreover, we hypothesise that the good performers will be able
to anticipate the robot's motion. Eye-tracking data can also be useful to verify
the anticipation hypothesis.

The experiment consists in two phases. The first phase is the observation phase
during which the participants watch the robot move and detect an obstacle in a
``playground''. The second phase requires the participants to interact with the
robot by placing an obstacle in its path to make it detect the obstacle before
reaching a certain target.  After each phase the participant explain the
behaviour of the robot. The sensor data was visualised through LEDs with three
visualisation conditions: {\sf TRUE} visualisation, {\sf RANDOM} visualisation and {\sf NONE}
visualisation. Through this experiment we investigate the following research
questions:

@@

The rest of the paper is organised as follows. The second section gives
a brief review of the eye-tracking research done to quantify the
expertise and performance. The third section presents the specificities
of the study. The fourth section lists the main results. The fifth
section discusses the results. Finally, the sixth section concludes the
paper.

@@

Our previous work also show a clear relation between gaze patterns and
task-based performance. In a pair-programming task, \cite{jermann2012effects}
showed that the good performing pairs have more synchronised gaze on different
parts of a program than the bad performing pairs. In a similar task,
\cite{sharma2012gaze} showed that the good performing pairs pay more attention
to the data-flow of the program than the poor performing pairs. Moreover,
\cite{sharma2013understanding} showed that while describing the functionality of
a program the well performing teams had more gaze on the variable modification
parts in the program while poor performing teams have equal distribution of gaze
on different parts of the program during similar phase of the task.  Thus we see
that previous research provides insights about the relationship between the gaze
patterns and the behavioural and task-based performance indicators in diverse
scenarios.

@@

In the experiment, there were 3 sensor data visualisation conditions.  For the
{\sf TRUE} visualisation condition, the number of illuminated LEDs, on the top of the
robot, is proportional to the intensity measured by the IR-sensor. When the
intensity of the reflected IR crosses a predefined threshold, the robot does a
90° turn to its right. For the two other conditions, the behaviour is the same,
simply the display changes. In the {\sf RANDOM} condition, the display shows as if the
robot was detecting the obstacle at random moments, while in the {\sf NONE}
visualization condition nothing is shown on top of the robot. The robot has a
fitting for a pencil that we used for the drawing of its trajectory.

@@

The playground (Figure~\ref{playground}) is designed in order to allow the
participants to get a reference of the previous behaviour of the robot, as well
as to allow us to localise the position of the robot, the position of the
obstacles and the position of the gaze on the video recorded from the
eye-tracker. The localisation is made possible by the use of unique fiducial
markers on the playground. On the right hand side, the observation phase takes
place and there will be the reference for the black and white obstacles. On the
left hand side, the interaction takes place.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{maze}
    \caption{\small Basic playground for the experiment. The fiducial markers
    are used for automatic localisation of the gaze pointers on the observation
    video.}

@@

One important fact to be considered is that the localisation of the
robot, the obstacle or the gaze pointer on the observation field for the
participant is not a trivial task. The fact that the participants are
free to move, poses a challenge to the eye-tracking analysis. The
movement of the participant causes change in the observation field of
the participant almost every frame of the eye-tracking video. To cope up
with the changing observation field we decided to put an array of
fiducial markers on the playground. Using the fiducial markers enabled
us to recreate the whole playground for every frame in the video
recorded from the eye-tracker's camera (Figure~\ref{expe}).

@@

\paragraph{Observation phase} The participant observes the Thymio II robot
approaching obstacles. The robot turns as soon as it detects the
obstacle. The participant watches the robot's behaviour for a black and a
white obstacle for 5 trials each. For the white, the robot turns
earlier; for the black obstacle the robot turns later. In the final
trial the robot is equipped with a pen to draw its trajectory for both
obstacles with different colours. After drawing, the participant is asked
the question: ``How does the robot work?''

\paragraph{Interaction phase} In the second phase, the participant is asked
to guide the robot to a goal, using a grey obstacle. The participant can put the
obstacle wherever he likes on the left hand side (Figure~\ref{playground}) of
the playground before the robot starts. We designed the grey obstacle to reflect
more infrared light (IR) than the black or the white obstacles therefore the
robot turns even earlier. This leads to a surprising effect. The participant has
only the robot's trajectories as a source of information. Each participant is
given 5 trials to put the obstacle in order to guide the robot to the goal. The
participant then has to answer the same question as in the observation phase,
response to which is considered as a final answer.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{setup}
    \caption{\small Experimental setup for observation phase. The observation
    video is recorded from the video camera in the front of the eye tracker (in
    the lop-left corner).}

@@

We transcribed the responses to perform a rating scheme on the level of
abstraction and correctness. A low level of abstraction consists in
describing the functionality of the robot in the terms of the electric
or programming aspects of the robot, a high level represents describing
the behaviour. For the levels of correctness, we categorised the
responses into correct or incorrect response. In TABLE I. we present
sample responses for the different categories.

\begin{table}[ht!]
    \centering
    \footnotesize
    \begin{tabular}{ccp{4.5cm}}
        \toprule
        Correct. & Abstract. & Sample answer \\
        \midrule
        \textit{incorrect} & \textit{low} & ``The robot turns at a given distance in respect to a colour, defined in the program.''\\ 
        \midrule
        \textit{incorrect} & \textit{high} & `Grey is the most dangerous, so the
        robot turn really early.''\\ 
        \midrule
        \textit{correct} & \textit{low} & ``The grey obstacle reflects a lot,
        the robot sends light. At a certain threshold the robot turns.'' \\ 
        \midrule
        \textit{correct} & \textit{high} & ``The grey obstacle is shinier, so
        more reflective, and better detectable.''\\ 
        \bottomrule
    \end{tabular}
    \caption{Sample answers for all classes}

@@

    \item The participant identified a wrong cause for the behaviour of the robot:
        ``The robot notices the tags on the side''.

    \item The participant held the code responsible for the cause of different
        behaviour: ``The robot turns at a given distance in respect to a color,
        defined in the program''.

    \item The participant tried to explain by analogies or did not explain the
        robots behaviour at an understandable level: ``Grey is the most
        dangerous, so the robot turns really early''.

@@

The most common form of eye-tracking data aggregation are the fixations.
Fixations depict the periods in the user's observation when the user
attends a relatively small part of the visual stimulus for a relatively
longer period of time. The various fixation detection algorithms are
summarised in~\cite{duchowski2007eye}. Following are the two gaze measures, derived
from the fixations, which we used in the present work.

@@

Average fixation on the reference side: The main area (right hand side
of figure 2) in the observation phase is treated as the reference side
in the interaction phase. The rationale behind keeping the lines drawn
by the robot in the observation phase is for the participants to be able
to refer to the prior behaviour of the robot in the interaction phase. We
measured the amount of time a participant looks at the reference side
and averaged this duration for the number of fixations on the reference
side for the particular participant.

@@

We found no bias for age, gender, expertise, student status or condition
in respect to the correctness, abstraction and type of mistake made for
the final answer. Surprisingly there is no correlation between the
visualisation condition and the correctness or abstraction of the
answers given. We could not find any anticipation patterns in the
eye-tracking data. We found some other relation between the different
measures. We present the results in following subsections.

@@

The average fixation duration on the robot during observation phase
(Figure~\ref{res1}) is significantly more in {\sf TRUE} and {\sf RANDOM} condition than in
the {\sf NONE} condition ($F[2,49]=3.68$, p = .03). This depicts that the
sensor data visualisation has an effect on the users' attention.

@@

The fact that the participants in the {\sf RANDOM} or {\sf NONE} visualisation
condition improved more than those with {\sf TRUE} condition is surprising
(Figure 5). There are two plausible explanations. The first is that
participants who see the robot in the {\sf TRUE} condition have a stronger
belief that the robot always behaves the same. They see the LEDs of the
robot turning on and therefore the robot still works. The ones with {\sf NONE}
visualisation do not have an indication whether the robot works the same
way as in the observation phase, and start experimenting earlier. The
second is that the participants in {\sf TRUE} condition just put the obstacles
out of detection range, because they could see different display
information and concluded therefore it may behave differently.

The fact that participants have lower average fixation duration on the
reference side in the interaction phase in the {\sf TRUE} condition than the
other two conditions actually supports the explanation that participants
with {\sf TRUE} condition had a higher initial belief in their cognitive model
of how does the robot work (Figure 6). They did not need to look onto
the reference given in observation phase. This also verifies our working
hypothesis that a carefully designed mapping between the robot's
behavior and its visual actuators can help in maximising the learning
effect during the interaction.

Another plausible explanation for the participants in {\sf NONE} and {\sf RANDOM} condition
being better performers than the {\sf TRUE} condition during the interaction phase
comes from the transparency of the robot's behaviour. The robot's
behaviour was not at all transparent in the former two conditions. This might
have caused the participants to put more effort in understanding the behaviour.
Therefore, they performed better.

@@

The advantages of mobile eye-tracking come with an additional cost. As
we mentioned earlier, the automatic localisation of the gaze pointer on
the visual stimulus is not a trivial task. To do this we have to use a
number of fiducial markers to the playground (Figure 2). We observed
that the fiducial markers cause a distraction to the participants. As a
future work, we are trying to find a suitable replacement to the
fiducial markers.

@@

We addressed two research questions in the beginning of the present
work. The results show that the way the sensor data is visualised has an
impact on both the gaze patterns and the performance (Question 2).
However, for the relation between the gaze patterns of the participants
and their performance outcomes (Question 1) we cannot claim a direct
relation. We found a negative correlation between the gaze-patterns and
the first improvement during the interaction phase. We could not find a
relation between the explanation of the participants and their
gaze-patterns. Moreover, we did not find any anticipation patterns in
the gaze data. The lack of evidence for the relation between
participants' explanations and their gaze patterns prevents us to make
any strong claims about the relation between the overall performance and
the gaze patterns.

The fact that we cannot find a relation between the gaze patterns and
all the factors in the performance might have its roots in the way we
designed the interaction phase experiment. As a working hypothesis we
said that there should be clear and carefully designed mapping between
the robots' behaviour and the sensor data visualization. Its environment
also affects the behaviour of the robot and hence the environment also
should be taken into consideration while designing the
behavior-visualisation mapping. We introduced a novelty (the gray
obstacle being more reflective) in the environment during the
interaction phase of the experiment that forced the participants to
revise their hypothesis and hence they failed to understand the robot's
functionality. This also serves as a word of caution in future while
designing such interactions. The observation phase serves as a training
session for the participants, which can later be tested during the
interaction phase. Any novel behaviour modality (in our case, the more
reflective grey obstacle) can have a negative impact on the
participants' understanding.
